---
title: "Untitled"
author: "Mathias Steilen"
date: "2024-04-25"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(tidymodels)
library(doParallel)
library(ggsci)
library(scales)
library(vip)

setwd(dirname(rstudioapi::getActiveDocumentContext()$path))

# Default theme for charts
theme_set(
  theme_bw() +
    theme(  
      plot.title = element_text(face = "bold", size = 14),
      plot.subtitle = element_text(
        face = "italic", size = 10, colour = "green50"
      )
    )
)
```

```{r}
df = read_csv("../../00 Data Retrieval and Cleaning/0_df_final_imputed_shifting_a.csv")
```

### Drop columns that can't be used

```{r}
# df = df |> 
#   select(-c(
#     auction_price_ch_de, allocatedCapacity_ch_de
#   ))
```


### Shift the other target variables

Anything that is a target variable in our project has not been shifted yet.

Anything that is not a target variable in some project is shifted according to the A, B, C logic that we discussed.

```{r}
df = df |> 
  mutate(
    # auction_price_ch_de = lag(auction_price_ch_de, n = 24),
    auction_price_de_ch = lag(auction_price_de_ch, n = 24),
    allocatedCapacity_ch_de = lag(allocatedCapacity_ch_de, n = 24),
    allocatedCapacity_de_ch = lag(allocatedCapacity_de_ch, n = 24),
    day_ahead_price_at = lag(day_ahead_price_at, n = 24),
    day_ahead_price_ch = lag(day_ahead_price_ch, n = 24),
    day_ahead_price_de = lag(day_ahead_price_de, n = 24),
    day_ahead_price_fr = lag(day_ahead_price_fr, n = 24),
    day_ahead_price_it = lag(day_ahead_price_it, n = 24),
  ) |> 
  drop_na()

df |> glimpse()
```

### Tuning Loop

```{r}
min_stamp = c(
  min(df$date), 
  ymd("2021-06-01"), 
  ymd("2023-03-01"),
  min(df$date)
)

max_stamp = c(
  ymd("2021-06-01"),
  ymd("2023-03-01"),
  max(df$date),
  max(df$date)
)
```

```{r}
for (i in 1:length(min_stamp)){
  
  # train and test split
  df_full = df |> 
    filter(date >= min_stamp[i]) |> 
    filter(date < max_stamp[i]) |> 
    arrange(date)
  
  train_ratio = 0.7
  train_index = ceiling(nrow(df_full)*train_ratio)
  test_index = floor(nrow(df_full)*(1-train_ratio))
  
  dt_train = df_full |> head(train_index)
  dt_test = df_full |> tail(test_index)
  
  # Time Series Split for tuning
  initial_split = 0.5
  increment = 0.1
  train_perc = 0.7
  
  initial_train = floor(nrow(dt_train)*(initial_split+increment)*train_perc)
  initial_test = round(floor(nrow(dt_train) * initial_split * (1-train_perc)))
  
  folds = rolling_origin(
    dt_train, 
    initial = initial_train, 
    assess = initial_test,
    skip = round(increment * nrow(dt_train)),
    cumulative = T
  )
  
  # Recipe for preprocessing
  lasso_rec <- recipe(auction_price_ch_de ~ ., data = dt_train) |>
    update_role(date, new_role = "ID") |>
    step_lag(all_predictors(), lag = c(24, 48, 72, 7*24, 14*24)) |> 
    step_zv(all_predictors()) |> 
    step_impute_mean(all_predictors()) |> 
    step_center(all_predictors()) |> 
    step_scale(all_predictors())
  
  
  # Model specification
  lasso_spec <- linear_reg(penalty = tune(), mixture = 1) |>
    set_engine("glmnet") |>
    set_mode("regression")
  
  lasso_wflow <- workflow() |> 
    add_recipe(lasso_rec) |> 
    add_model(lasso_spec)
  
  # Setting grid for hyperparameter tuning
  lasso_grid = crossing(
    penalty = seq(0.0001, 10, length.out = 500)
  )
  
  start_time = Sys.time()
  # Tuning Hyperparameters
  unregister_dopar <- function() {
    env <- foreach:::.foreachGlobals
    rm(list=ls(name=env), pos=env)
  }
  
  cl <- makePSOCKcluster(6)
  registerDoParallel(cl)
  
  lasso_tune <- tune_grid(object = lasso_wflow,
                          grid = lasso_grid,
                          resamples = folds)
  
  stopCluster(cl)
  unregister_dopar()
  end_time = Sys.time()
  print(end_time - start_time)
  
  # Write results to csv
  tuning_results = lasso_tune |> 
    collect_metrics()
  
  tuning_results |> 
    write_csv(paste0("tuning_grid_results_split_", i, ".csv"))
  
  # Fit the best model
  best_combination = tuning_results |> 
    filter(.metric == "rmse") |> 
    arrange(mean) |> 
    head(1)
  
  lasso_spec <- linear_reg(penalty = best_combination$penalty,
                           mixture = 1) |>
    set_engine("glmnet") |>
    set_mode("regression")
  
  lasso_wflow <- workflow() |> 
    add_recipe(lasso_rec) |> 
    add_model(lasso_spec)
  
  lasso_final_fit <- lasso_wflow |> 
    fit(dt_train)
  
  # Write non-zero coefficients
  lasso_final_fit |> 
    tidy() |> 
    filter(abs(estimate) >= 1e-10) |> 
    write_csv(paste0("non_zero_coefficients_split_", i, ".csv"))
  
  # Write predictions
  lasso_final_fit |> 
    augment(dt_test) |> 
    select(date, auction_price_ch_de, .pred) |> 
    write_csv(paste0("holdout_predictions_split_", i, ".csv"))
  
}
```

### Model Evaluation

TBD: 
- Read the tuning results
- Read the fit results
- Calc holdout metrics
- Plots for final presentation

```{r}
list.files()
```

```{r}
holdout_preds = bind_rows(
  read_csv("holdout_predictions_split_1.csv") |> mutate(period = "Split 1"),
  read_csv("holdout_predictions_split_2.csv") |> mutate(period = "Split 2"),
  read_csv("holdout_predictions_split_3.csv") |> mutate(period = "Split 3"),
)

eval_metrics = metric_set(rsq, rmse, mae, mape)

holdout_preds |> 
  group_by(period) |> 
  eval_metrics(truth = .pred, estimate = day_ahead_price_ch) |> 
  select(-.estimator) |> 
  pivot_wider(names_from = .metric, values_from = .estimate)
  

holdout_preds |> 
  ggplot(aes(day_ahead_price_ch , .pred)) +
  geom_point(alpha = 0.2, colour = "midnightblue", size = 2) +
  geom_abline(lty = "dashed", colour = "grey50") +
  scale_x_continuous(labels = scales::comma_format()) +
  scale_y_continuous(labels = scales::comma_format()) +
  facet_wrap(~ period, scales = "free") +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5))

holdout_preds |> 
  pivot_longer(-c(date, period)) |> 
  ggplot(aes(date, value, colour = name)) +
  geom_line(alpha = 0.75, linewidth = 0.75) +
  facet_wrap(~ period, ncol = 1, scales = "free") +
  labs(title = "Out-Of-Sample Time Series",
       y = NULL,
       x = NULL,
       colour = NULL) +
  ggsci::scale_colour_jama() +
  scale_y_continuous(labels = scales::comma_format(suffix = " MWh")) +
  theme_bw() +
  theme(plot.title = element_text(size = 14, face = "bold"),
        plot.subtitle = element_text(face = "italic", size = 12,
                                     colour = "grey50"),
        legend.position = "bottom")
```
