---
title: "Untitled"
author: "Mathias Steilen"
date: "2024-04-25"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(tidymodels)
library(doParallel)
library(ggsci)
library(scales)
library(vip)

setwd(dirname(rstudioapi::getActiveDocumentContext()$path))

# Default theme for charts
theme_set(
  theme_bw() +
    theme(  
      plot.title = element_text(face = "bold", size = 14),
      plot.subtitle = element_text(
        face = "italic", size = 10, colour = "green50"
      )
    )
)
```

```{r}
df = read_csv("../../00 Data Retrieval and Cleaning/0_df_final_imputed_shifting_a.csv")
```

### Drop columns that can't be used

```{r}
# df = df |> 
#   select(-c(
#     auction_price_ch_de, allocatedCapacity_ch_de
#   ))
```


### Shift the other target variables

Anything that is a target variable in our project has not been shifted yet.

Anything that is not a target variable in some project is shifted according to the A, B, C logic that we discussed.

```{r}
df = df |> 
  mutate(
    auction_price_ch_de = lag(auction_price_ch_de, n = 24),
    auction_price_de_ch = lag(auction_price_de_ch, n = 24),
    allocatedCapacity_ch_de = lag(allocatedCapacity_ch_de, n = 24),
    allocatedCapacity_de_ch = lag(allocatedCapacity_de_ch, n = 24),
    day_ahead_price_at = lag(day_ahead_price_at, n = 24),
    # day_ahead_price_ch = lag(day_ahead_price_ch, n = 24),
    day_ahead_price_de = lag(day_ahead_price_de, n = 24),
    day_ahead_price_fr = lag(day_ahead_price_fr, n = 24),
    day_ahead_price_it = lag(day_ahead_price_it, n = 24),
  ) |> 
  drop_na()

df |> glimpse()
```

### Tuning Loop

```{r}
# train and test split
dt_train = df |> 
  filter(date >= ymd("2023-01-01")) |> 
  filter(date < ymd("2024-01-01"))

dt_test = df |> 
  filter(date >= ymd("2024-01-01"))

# Time Series Split for tuning
initial_split = 0.5
increment = 0.1
train_perc = 0.7

initial_train = floor(nrow(dt_train)*(initial_split+increment)*train_perc)
initial_test = round(floor(nrow(dt_train) * initial_split * (1-train_perc)))

folds = rolling_origin(
  dt_train, 
  initial = initial_train, 
  assess = initial_test,
  skip = round(increment * nrow(dt_train)),
  cumulative = T
)

# Recipe for preprocessing
xg_rec <- recipe(day_ahead_price_ch ~ ., data = dt_train) |>
  update_role(date, new_role = "ID") |>
  step_zv(all_predictors()) |> 
  step_impute_mean(all_predictors()) |> 
  step_center(all_predictors()) |> 
  step_scale(all_predictors())

# Model specification
xg_spec <- boost_tree(mtry = tune(),
                      trees = tune(),
                      min_n = tune(),
                      tree_depth = tune(),
                      learn_rate = tune()) |>
  set_engine("xgboost") |>
  set_mode("regression")

xg_wflow <- workflow() |> 
  add_recipe(xg_rec) |> 
  add_model(xg_spec)

# Setting grid for hyperparameter tuning
xg_grid <- grid_latin_hypercube(
  finalize(mtry(), dt_train),
  trees(),
  min_n(),
  tree_depth(),
  learn_rate(),
  size = 100
) # 6.2 hours with 100

start_time = Sys.time()
# Tuning Hyperparameters
unregister_dopar <- function() {
  env <- foreach:::.foreachGlobals
  rm(list=ls(name=env), pos=env)
}

cl <- makePSOCKcluster(6)
registerDoParallel(cl)

xg_tune <- tune_grid(object = xg_wflow,
                     grid = xg_grid,
                     resamples = folds)

stopCluster(cl)
unregister_dopar()
end_time = Sys.time()
print(end_time - start_time)

# Write results to csv
tuning_results = xg_tune |> 
  collect_metrics()

tuning_results |> 
  write_csv("tuning_grid_results_split.csv")

# Fit the best model
best_combination = tuning_results |> 
  filter(.metric == "rmse") |> 
  arrange(mean) |> 
  head(1)

xg_spec <- boost_tree(
  mtry = best_combination$mtry,
  trees = best_combination$trees,
  min_n = best_combination$min_n,
  tree_depth = best_combination$tree_depth,
  learn_rate = best_combination$learn_rate,
) |>
  set_engine("xgboost") |>
  set_mode("regression")

xg_wflow <- workflow() |> 
  add_recipe(xg_rec) |> 
  add_model(xg_spec)

xg_final_fit <- xg_wflow |> 
  fit(dt_train)

# Write variable importance
xg_final_fit |>
  extract_fit_parsnip() |>
  vip::vi() |> 
  write_csv("variable_importance_split.csv")

# Write predictions
xg_final_fit |> 
  augment(dt_test) |> 
  select(date, day_ahead_price_ch, .pred) |> 
  write_csv("holdout_predictions_split.csv")
```


### Model Evaluation

TBD: 
- Read the tuning results
- Read the fit results
- Calc holdout metrics
- Plots for final presentation

```{r}
holdout_preds = bind_rows(
  read_csv("holdout_predictions_split_1.csv") |> mutate(period = "Split 1"),
  read_csv("holdout_predictions_split_2.csv") |> mutate(period = "Split 2"),
  read_csv("holdout_predictions_split_3.csv") |> mutate(period = "Split 3"),
)

eval_metrics = metric_set(rsq, rmse, mae, mape)

holdout_preds |> 
  group_by(period) |> 
  eval_metrics(truth = .pred, estimate = day_ahead_price_ch) |> 
  select(-.estimator) |> 
  pivot_wider(names_from = .metric, values_from = .estimate)


holdout_preds |> 
  ggplot(aes(day_ahead_price_ch , .pred)) +
  geom_point(alpha = 0.2, colour = "midnightblue", size = 2) +
  geom_abline(lty = "dashed", colour = "grey50") +
  scale_x_continuous(labels = scales::comma_format()) +
  scale_y_continuous(labels = scales::comma_format()) +
  facet_wrap(~ period, scales = "free") +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5))

holdout_preds |> 
  pivot_longer(-c(date, period)) |> 
  ggplot(aes(date, value, colour = name)) +
  geom_line(alpha = 0.75, linewidth = 0.75) +
  facet_wrap(~ period, ncol = 1, scales = "free") +
  labs(title = "Out-Of-Sample Time Series",
       y = NULL,
       x = NULL,
       colour = NULL) +
  ggsci::scale_colour_jama() +
  scale_y_continuous(labels = scales::comma_format(suffix = " MWh")) +
  theme_bw() +
  theme(plot.title = element_text(size = 14, face = "bold"),
        plot.subtitle = element_text(face = "italic", size = 12,
                                     colour = "grey50"),
        legend.position = "bottom")
```
