---
title: "STUDENT T"
author: "Olivier Zehnder"
date: "2024-05-22"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
### Modelling of Swiss Day ahead prices

# Load necessary libraries

library(rugarch)
library(forecast)
library(dplyr)

#Functions
# Define a function to calculate R-squared
calculate_r_squared <- function(true_values, fitted_values) {
  # Calculate the mean of true values
  true_mean <- mean(true_values)
  
  # Calculate total sum of squares (TSS)
  TSS <- sum((true_values - true_mean)^2)
  
  # Calculate residual sum of squares (RSS)
  RSS <- sum((true_values - fitted_values)^2)
  
  # Calculate R-squared
  R_squared <- 1 - (RSS / TSS)
  
  return(R_squared)
}
# Load your dataset

```

Loading data:
```{r}
df_CH_DA <- read.csv("~/Downloads/0_df_predictive_ch_spot_price.csv", header=TRUE)
df_CH_DA$date <- as.Date(df_CH_DA$date)
df_CH_DA$cal_day_in_month = as.factor(df_CH_DA$cal_day_in_month)
df_CH_DA$cal_day_in_week = as.factor(df_CH_DA$cal_day_in_week)
df_CH_DA$cal_hour_in_day= as.factor(df_CH_DA$cal_hour_in_day)
df_CH_DA$cal_month = as.factor(df_CH_DA$cal_month)
df_CH_DA$cal_quarter = as.factor(df_CH_DA$cal_quarter)


df_DE_DA <- read.csv("~/Downloads/0_df_predictive_chde_auction_price.csv", header=TRUE)
df_DE_DA$date <- as.Date(df_DE_DA$date)
df_DE_DA$cal_day_in_month = as.factor(df_DE_DA$cal_day_in_month)
df_DE_DA$cal_day_in_week = as.factor(df_DE_DA$cal_day_in_week)
df_DE_DA$cal_hour_in_day= as.factor(df_DE_DA$cal_hour_in_day)
df_DE_DA$cal_month = as.factor(df_DE_DA$cal_month)
df_DE_DA$cal_quarter = as.factor(df_DE_DA$cal_quarter)

df_DE_CH_AUCT <-read.csv("~/Downloads/0_df_predictive_dech_auction_price.csv", header=TRUE)
df_DE_CH_AUCT$date <- as.Date(df_DE_CH_AUCT$date)
df_DE_CH_AUCT$cal_day_in_month = as.factor(df_DE_CH_AUCT$cal_day_in_month)
df_DE_CH_AUCT$cal_day_in_week = as.factor(df_DE_CH_AUCT$cal_day_in_week)
df_DE_CH_AUCT$cal_hour_in_day= as.factor(df_DE_CH_AUCT$cal_hour_in_day)
df_DE_CH_AUCT$cal_month = as.factor(df_DE_CH_AUCT$cal_month)
df_DE_CH_AUCT$cal_quarter = as.factor(df_DE_CH_AUCT$cal_quarter)

df_CH_DE_AUCT <- read.csv("~/Downloads/0_df_predictive_chde_auction_price.csv", header=TRUE)
df_CH_DE_AUCT$date <- as.Date(df_CH_DE_AUCT$date)
df_CH_DE_AUCT$cal_day_in_month = as.factor(df_CH_DE_AUCT$cal_day_in_month)
df_CH_DE_AUCT$cal_day_in_week = as.factor(df_CH_DE_AUCT$cal_day_in_week)
df_CH_DE_AUCT$cal_hour_in_day= as.factor(df_CH_DE_AUCT$cal_hour_in_day)
df_CH_DE_AUCT$cal_month = as.factor(df_CH_DE_AUCT$cal_month)
df_CH_DE_AUCT$cal_quarter = as.factor(df_CH_DE_AUCT$cal_quarter)

```

Getting rid of linear dependent cols:
```{r}
df_CH_DA_train= arrange(df_CH_DA, date)[1:8734,]
fit_cleaning= lm(day_ahead_price_ch~., data=df_CH_DA_train)

# Identify predictors with NA coefficients (indicating rank deficiency)
na_predictors <- names(fit_cleaning$coefficients)[is.na(fit_cleaning$coefficients)]

df_CH_DA <- df_CH_DA %>%
  select(-any_of(na_predictors))

df_DE_DA <- df_DE_DA %>%
  select(-any_of(na_predictors))

df_DE_CH_AUCT <- df_DE_CH_AUCT %>%
  select(-any_of(na_predictors))

df_CH_DE_AUCT <- df_CH_DE_AUCT %>%
  select(-any_of(na_predictors))
```



Split the data into training and testing data. Training data is data from the year 2023, testing data from the year 2024
```{r}


df_CH_DA_train= arrange(df_CH_DA, date)[1:8734,] #Split data at end of 2023 / Beginning of 2024  into train and test
df_CH_DA_test= arrange(df_CH_DA, date)[8735:9477,]

df_DE_DA_train= arrange(df_DE_DA, date)[1:8734,]
df_DE_DA_test= arrange(df_DE_DA, date)[8735:9477,]

df_DE_CH_AUCT_train= arrange(df_DE_CH_AUCT, date)[1:8734,]
df_DE_CH_AUCT_test= arrange(df_DE_CH_AUCT, date)[8735:9477,]

df_CH_DE_AUCT_train= arrange(df_CH_DE_AUCT, date)[1:8734,]
df_CH_DE_AUCT_test= arrange(df_CH_DE_AUCT, date)[8735:9477,]

rm(df_CH_DA)
rm(df_DE_DA)
rm(df_DE_CH_AUCT)
rm(df_CH_DE_AUCT)
```


```{r}
df_CH_DA_train_numeric <- as.data.frame(lapply(df_CH_DA_train, function(x) as.numeric(as.character(x))))
df_DE_DA_train_numeric <- as.data.frame(lapply(df_DE_DA_train, function(x) as.numeric(as.character(x))))
df_DE_CH_AUCT_train_numeric <- as.data.frame(lapply(df_DE_CH_AUCT_train, function(x) as.numeric(as.character(x))))
df_CH_DE_AUCT_train_numeric <- as.data.frame(lapply(df_CH_DE_AUCT_train, function(x) as.numeric(as.character(x))))


df_CH_DA_test_numeric <- as.data.frame(lapply(df_CH_DA_test, function(x) as.numeric(as.character(x))))
df_DE_DA_test_numeric <- as.data.frame(lapply(df_DE_DA_test, function(x) as.numeric(as.character(x))))
df_DE_CH_AUCT_test_numeric <- as.data.frame(lapply(df_DE_CH_AUCT_test, function(x) as.numeric(as.character(x))))
df_CH_DE_AUCT_test_numeric <- as.data.frame(lapply(df_CH_DE_AUCT_test, function(x) as.numeric(as.character(x))))

xreg_train_CH_DA <- select(df_CH_DA_train_numeric, -c("day_ahead_price_ch", "date")) %>%
  as.matrix()
xreg_train_DE_DA <- select(df_DE_DA_train_numeric, -c("day_ahead_price_de", "date"))%>%
  as.matrix()
xreg_train_DE_CH_AUCT <- select(df_DE_CH_AUCT_train_numeric, -c("auction_price_de_ch", "date")) %>%
  as.matrix()
xreg_train_CH_DE_AUCT <- select(df_CH_DE_AUCT_train_numeric, -c("auction_price_ch_de", "date")) %>%
  as.matrix()

# rank of the matrix to invert in ARIMA:

pivotal_columns <- function(X){
  rank<- qr(t(X) %*% X)$rank
  n <- ncol(X)
  pivotal_columns <- numeric(0)
  
  for (i in 1:n) {
    # Include the current column in the test matrix
    test_matrix <- X[,-i]
    
    # Check if the rank remains 173
    if (qr(t(test_matrix)%*%test_matrix)$rank != rank) {
      pivotal_columns <- c(pivotal_columns, i)
    }
  }
  return(pivotal_columns)
}

#pivot_col_CH_DA <- pivotal_columns(xreg_train_CH_DA)
#pivot_col_DE_DA <- pivotal_columns(xreg_train_DE_DA)
#pivot_col_DE_CH_AUCT <- pivotal_columns(xreg_train_DE_CH_AUCT)
#ot_col_CH_DE_AUCT <- pivotal_columns(xreg_train_CH_DE_AUCT)
```

```{r}
xreg_train_CH_DA<- select(df_CH_DA_train_numeric, -c("day_ahead_price_ch", "date"))
xreg_train_DE_DA<- select(df_DE_DA_train_numeric, -c("day_ahead_price_de", "date"))
xreg_train_DE_CH_AUCT<- select(df_DE_CH_AUCT_train_numeric, -c("auction_price_de_ch", "date"))
xreg_train_CH_DE_AUCT<- select(df_CH_DE_AUCT_train_numeric, -c("auction_price_ch_de", "date"))

xreg_test_CH_DA<- select(df_CH_DA_test_numeric, -c("day_ahead_price_ch", "date"))
xreg_test_DE_DA<- select(df_DE_DA_test_numeric, -c("day_ahead_price_de", "date"))
xreg_test_DE_CH_AUCT<- select(df_DE_CH_AUCT_test_numeric, -c("auction_price_de_ch", "date"))
xreg_test_CH_DE_AUCT<- select(df_CH_DE_AUCT_test_numeric, -c("auction_price_ch_de", "date"))

CH_DA_price= df_CH_DA_train$day_ahead_price_ch
DE_DA_price= df_DE_DA_train$day_ahead_price_de
AUCT_CH_DE_price= df_CH_DE_AUCT_train$auction_price_ch_de
AUCT_DE_CH_price= df_DE_CH_AUCT_train$auction_price_de_ch

```



#retraining
```{r}
# Required Libraries
library(rugarch)

# Function to calculate R-squared
calculate_r_squared <- function(actual, predicted) {
  ss_res <- sum((actual - predicted) ^ 2)
  ss_tot <- sum((actual - mean(actual)) ^ 2)
  r2 <- 1 - (ss_res / ss_tot)
  return(r2)
}

# Initialize variables
predictions_DA_CH <- numeric()
predictions_DA_DE <- numeric()
predictions_AUCT_CH_DE <- numeric()
predictions_AUCT_DE_CH <- numeric()

#Chunk size for predictions
chunk_size <- 24

predictors_DA_CH = c("auction_price_ch_de", "auction_price_de_ch", "day_ahead_price_at","day_ahead_price_de" ,"day_ahead_price_fr","day_ahead_price_it", "actual_load_at", "actual_load_ch","actual_load_de","allocated_capacity_ch_de","allocated_capacity_de_ch", "cal_day_in_month",  "cal_day_in_week","cal_hour_in_day","cal_quarter" )

predictors_DA_DE = c("auction_price_ch_de", "auction_price_de_ch", "day_ahead_price_at","day_ahead_price_ch" ,"day_ahead_price_fr","day_ahead_price_it", "actual_load_at", "actual_load_ch","actual_load_de","allocated_capacity_ch_de","allocated_capacity_de_ch", "cal_day_in_month",  "cal_day_in_week","cal_hour_in_day","cal_quarter" )

predictors_AUCT_CH_DE = c("auction_price_de_ch", "day_ahead_price_at","day_ahead_price_ch","day_ahead_price_de" ,"day_ahead_price_fr","day_ahead_price_it", "actual_load_at", "actual_load_ch", "actual_load_de","allocated_capacity_ch_de","allocated_capacity_de_ch", "cal_day_in_month",  "cal_day_in_week","cal_hour_in_day","cal_quarter" )

predictors_AUCT_DE_CH = c("auction_price_ch_de", "day_ahead_price_at","day_ahead_price_ch","day_ahead_price_de" ,"day_ahead_price_fr","day_ahead_price_it", "actual_load_at", "actual_load_ch", "actual_load_de","allocated_capacity_ch_de","allocated_capacity_de_ch", "cal_day_in_month",  "cal_day_in_week","cal_hour_in_day","cal_quarter")

#Ensure the initial training data is in the correct format
train_data_CH_DA <- cbind(asinh(df_CH_DA_train$day_ahead_price_ch), xreg_train_CH_DA[,predictors_DA_CH])
colnames(train_data_CH_DA)[1] <- "asinh_price_ch"
train_data_DE_DA <- cbind(asinh(df_DE_DA_train$day_ahead_price_de), xreg_train_DE_DA[,predictors_DA_DE])
colnames(train_data_DE_DA)[1] <- "asinh_price_de"
train_data_CH_DE_AUCT <- cbind(asinh(df_CH_DE_AUCT_train$auction_price_ch_de), xreg_train_CH_DE_AUCT[, predictors_AUCT_CH_DE])
colnames(train_data_CH_DE_AUCT)[1] <- "asinh_price_ch_de"
train_data_DE_CH_AUCT <- cbind(asinh(df_DE_CH_AUCT_train$auction_price_de_ch), xreg_train_DE_CH_AUCT[, predictors_AUCT_DE_CH])
colnames(train_data_DE_CH_AUCT)[1] <- "asinh_price_de_ch"

  
  garch_model_DA_CH <- ugarchspec(mean.model = list(armaOrder = c(2,2), include.mean = TRUE, external.regressors = as.matrix(select(train_data_CH_DA,-"asinh_price_ch"))), variance.model = list(model = "sGARCH",garchOrder = c(1, 1), external.regressors = as.matrix(select(train_data_CH_DA,-"asinh_price_ch"))), distribution.model = "std")
  
  garch_model_DA_DE <- ugarchspec(mean.model = list(armaOrder = c(2,2), include.mean = TRUE,external.regressors = as.matrix(select(train_data_DE_DA, -"asinh_price_de"))), variance.model = list(model = "sGARCH", garchOrder = c(1, 1),  external.regressors = as.matrix(select(train_data_DE_DA, -"asinh_price_de"))), distribution.model = "std")

garch_model_CH_DE_AUCT <- ugarchspec(mean.model = list(armaOrder = c(2,2), include.mean = TRUE, external.regressors = as.matrix(select(train_data_CH_DE_AUCT, -"asinh_price_ch_de"))), variance.model = list(model = "sGARCH",garchOrder = c(1, 1), external.regressors = as.matrix(select(train_data_CH_DE_AUCT, -"asinh_price_ch_de"))), distribution.model = "std")

garch_model_DE_CH_AUCT <- ugarchspec(mean.model = list(armaOrder = c(2,2), include.mean = TRUE, external.regressors = as.matrix(select(train_data_DE_CH_AUCT, -"asinh_price_de_ch"))), variance.model = list(model = "sGARCH", garchOrder = c(1, 1), external.regressors = as.matrix(select(train_data_DE_CH_AUCT, -"asinh_price_de_ch"))), distribution.model = "std")

  
fit_DA_CH_ARMA_GARCH <- ugarchfit(spec = garch_model_DA_CH, data = train_data_CH_DA, solver="hybrid")
fit_DA_DE_ARMA_GARCH <- ugarchfit(spec = garch_model_DA_DE,data =  train_data_DE_DA, solver="hybrid")
fit_AUCT_CH_DE_ARMA_GARCH <- ugarchfit(spec = garch_model_CH_DE_AUCT, data =  train_data_CH_DE_AUCT, solver="hybrid")
fit_AUCT_DE_CH_ARMA_GARCH <- ugarchfit(spec = garch_model_DE_CH_AUCT, data = train_data_DE_CH_AUCT, solver="hybrid")

start_pars_DA_CH <- coef(fit_DA_CH_ARMA_GARCH)
start_pars_DA_DE <- coef(fit_DA_DE_ARMA_GARCH)
start_pars_AUCT_CH_DE<- coef(fit_AUCT_CH_DE_ARMA_GARCH)
start_pars_AUCT_DE_CH <- coef(fit_AUCT_DE_CH_ARMA_GARCH)



# Rolling Forecast with Retraining
for (i in seq(1, nrow(df_CH_DA_test), by = chunk_size)) {
  end <- min(chunk_size, nrow(df_CH_DA_test)-i+1)
  print(paste("Processing chunk:", i, "to", end))
  
  garch_model_DA_CH <- ugarchspec(mean.model = list(armaOrder = c(2,2), include.mean = TRUE, external.regressors = as.matrix(select(train_data_CH_DA,-c("asinh_price_ch")))), variance.model = list(model = "sGARCH",garchOrder = c(1, 1), external.regressors = as.matrix(select(train_data_CH_DA,-c("asinh_price_ch")))), distribution.model = "std")
  
  garch_model_DA_DE <- ugarchspec(mean.model = list(armaOrder = c(2,2), include.mean = TRUE,external.regressors = as.matrix(select(train_data_DE_DA, -c("asinh_price_de")))), variance.model = list(model = "sGARCH", garchOrder = c(1, 1),  external.regressors = as.matrix(select(train_data_DE_DA, -c("asinh_price_de")))), distribution.model = "std")

garch_model_CH_DE_AUCT <- ugarchspec(mean.model = list(armaOrder = c(2,2), include.mean = TRUE, external.regressors = as.matrix(select(train_data_CH_DE_AUCT, -c("asinh_price_ch_de")))), variance.model = list(model = "sGARCH",garchOrder = c(1, 1),external.regressors = as.matrix(select(train_data_CH_DE_AUCT, -c("asinh_price_ch_de")))), distribution.model = "sstd")

garch_model_DE_CH_AUCT <- ugarchspec(mean.model = list(armaOrder = c(2,2), include.mean = TRUE, external.regressors = as.matrix(select(train_data_DE_CH_AUCT, -c("asinh_price_de_ch")))), variance.model = list(model = "sGARCH", garchOrder = c(1, 1), external.regressors = as.matrix(select(train_data_DE_CH_AUCT, -c("asinh_price_de_ch")))), distribution.model = "sstd")

  
  fit_DA_CH_ARMA_GARCH <- ugarchfit(spec = garch_model_DA_CH, data = train_data_CH_DA, solver="hybrid", start.pars = start_pars_DA_CH)
  fit_DA_DE_ARMA_GARCH <- ugarchfit(spec = garch_model_DA_DE,data =  train_data_DE_DA, solver="hybrid",  start.pars =start_pars_DA_DE)
  fit_AUCT_CH_DE_ARMA_GARCH <- ugarchfit(spec = garch_model_CH_DE_AUCT, data =  train_data_CH_DE_AUCT, solver="hybrid", start.pars =start_pars_AUCT_CH_DE)
fit_AUCT_DE_CH_ARMA_GARCH <- ugarchfit(spec = garch_model_DE_CH_AUCT, data = train_data_DE_CH_AUCT, solver="hybrid", start.pars =start_pars_AUCT_DE_CH)
 print("check 1: " )

  
    predicted_values_DA_CH <- sinh(fitted(ugarchforecast(fit_DA_CH_ARMA_GARCH, n.ahead = end, external.forecasts = list(mregfor = as.matrix(xreg_test_CH_DA[i:(i+end-1),predictors_DA_CH])))))
    
    predicted_values_DA_DE <-  sinh(fitted(ugarchforecast(fit_DA_DE_ARMA_GARCH, n.ahead = end, external.forecasts = list(mregfor = as.matrix(xreg_test_DE_DA[i:(i+end-1),predictors_DA_DE])))))

predicted_values_AUCT_DE_CH <-  sinh(fitted(ugarchforecast(fit_AUCT_DE_CH_ARMA_GARCH, n.ahead = end, external.forecasts = list(mregfor = as.matrix(xreg_test_DE_CH_AUCT[i:(i+end-1),predictors_AUCT_DE_CH])))))

predicted_values_AUCT_CH_DE <-  sinh(fitted(ugarchforecast(fit_AUCT_CH_DE_ARMA_GARCH, n.ahead = end, external.forecasts = list(mregfor = as.matrix(xreg_test_CH_DE_AUCT[i:(i+end-1),predictors_AUCT_CH_DE])))))
  

  print("check 2: " )  
  # Store predictions
 predictions_DA_CH <- c(predictions_DA_CH, predicted_values_DA_CH)
 predictions_DA_DE <- c(predictions_DA_DE, predicted_values_DA_DE)
 predictions_AUCT_CH_DE <- c(predictions_AUCT_CH_DE, predicted_values_AUCT_CH_DE)
 predictions_AUCT_DE_CH <- c(predictions_AUCT_DE_CH, predicted_values_AUCT_DE_CH)
 
    print("check 3: " )
  # Update the training data with the new observations
  new_data_CH_DA <- cbind(asinh(df_CH_DA_test$day_ahead_price_ch[i:(end+i-1)]), xreg_test_CH_DA[i:(end+i-1),predictors_DA_CH])
  colnames(new_data_CH_DA)[1] <- "asinh_price_ch"
  
  new_data_DE_DA <- cbind(asinh(df_DE_DA_test$day_ahead_price_de[i:(end+i-1)]), xreg_test_DE_DA[i:(end+i-1),predictors_DA_DE])
  colnames(new_data_DE_DA)[1] <- "asinh_price_de"
  
  new_data_CH_DE_AUCT <- cbind(asinh(df_CH_DE_AUCT_test$auction_price_ch_de[i:(end+i-1)]), xreg_test_CH_DE_AUCT[i:(end+i-1),predictors_AUCT_CH_DE])
  colnames(new_data_CH_DE_AUCT)[1] <- "asinh_price_ch_de"
  
  new_data_DE_CH_AUCT <- cbind(asinh(df_DE_CH_AUCT_test$auction_price_de_ch[i:(end+i-1)]), xreg_test_DE_CH_AUCT[i:(end+i-1),predictors_AUCT_DE_CH] )
  colnames(new_data_DE_CH_AUCT)[1] <- "asinh_price_de_ch"

  
  
    print("check 4: " )
  train_data_CH_DA <- rbind(train_data_CH_DA, new_data_CH_DA)
  train_data_DE_DA <- rbind(train_data_DE_DA, new_data_DE_DA)
  train_data_CH_DE_AUCT <- rbind(train_data_CH_DE_AUCT, new_data_CH_DE_AUCT)
  train_data_DE_CH_AUCT <- rbind(train_data_DE_CH_AUCT, new_data_DE_CH_AUCT)
  
start_pars_DA_CH <- coef(fit_DA_CH_ARMA_GARCH)
start_pars_DA_DE <- coef(fit_DA_DE_ARMA_GARCH)
start_pars_AUCT_CH_DE<- coef(fit_AUCT_CH_DE_ARMA_GARCH)
start_pars_AUCT_DE_CH <- coef(fit_AUCT_DE_CH_ARMA_GARCH)
}

# Calculate R-squared and MSE for the entire test set
R2_DA_CH <- calculate_r_squared(df_CH_DA_test$day_ahead_price_ch[1:length(predictions_DA_CH)], predictions_DA_CH)
mse_DA_CH <- mean((df_CH_DA_test$day_ahead_price_ch[1:length(predictions_DA_CH)] - predictions_DA_CH)^2)

R2_DA_DE <- calculate_r_squared(df_DE_DA_test$day_ahead_price_de[1:length(predictions_DA_DE)], predictions_DA_DE)
mse_DA_DE <- mean((df_DE_DA_test$day_ahead_price_de[1:length(predictions_DA_DE)] - predictions_DA_DE)^2)

R2_AUCT_CH_DE <- calculate_r_squared(df_CH_DE_AUCT_test$auction_price_ch_de[1:length(predictions_AUCT_CH_DE)], pmax(0, predictions_AUCT_CH_DE))
mse_AUCT_CH_DE <- mean((df_CH_DE_AUCT_test$auction_price_ch_de[1:length(predictions_AUCT_CH_DE)] - pmax(0, predictions_AUCT_CH_DE))^2)

R2_AUCT_DE_CH <- calculate_r_squared(df_DE_CH_AUCT_test$auction_price_de_ch[1:length(predictions_AUCT_DE_CH)], pmax(0, predictions_AUCT_DE_CH))
mse_AUCT_DE_CH <- mean((df_DE_CH_AUCT_test$auction_price_de_ch[1:length(predictions_AUCT_DE_CH)] - pmax(0, predictions_AUCT_DE_CH))^2)

# Create a dataframe for the results
result_table <- data.frame(
  rmse = c(mse_DA_CH, mse_DA_DE, mse_AUCT_CH_DE, mse_AUCT_DE_CH),
  rsquared = c(R2_DA_CH, R2_DA_DE, R2_AUCT_CH_DE, R2_AUCT_DE_CH),
  row.names = c("Swiss DA Prices", "German DA Prices", "Auction Price CH-DE", "Auction Price DE-CH")
)

# Print the result
print(result_table)

```


```{r}

#Visualization of fit:
# Extract the fitted values
fitted_values_DA_CH <- sinh(fitted(fit_DA_CH_ARMA_GARCH))
fitted_values_DA_DE <- sinh(fitted(fit_DA_DE_ARMA_GARCH))
fitted_values_AUCT_CH_DE <- sinh(fitted(fit_AUCT_CH_DE_ARMA_GARCH))
fitted_values_AUCT_DE_CH <- sinh(fitted(fit_AUCT_DE_CH_ARMA_GARCH))



# Combine the original and fitted values into data frames
df_fitted_DA_CH <- data.frame(Time = 7734:8734, Original = df_CH_DA_train$day_ahead_price_ch[7734:8734], Fitted = fitted_values_DA_CH[7734:8734])
df_fitted_DA_DE <- data.frame(Time = 7734:8734, Original = df_DE_DA_train$day_ahead_price_de[7734:8734], Fitted = fitted_values_DA_DE[7734:8734])
df_fitted_AUCT_CH_DE <- data.frame(Time=7734:8734, Original = df_CH_DE_AUCT_train$auction_price_ch_de[7734:8734], Fitted = fitted_values_AUCT_CH_DE[7734:8734])
df_fitted_AUCT_DE_CH <- data.frame(Time=7734:8734, Original = df_DE_CH_AUCT_train$auction_price_de_ch[7734:8734], Fitted = fitted_values_AUCT_DE_CH[7734:8734])
transparent_firebrick <- adjustcolor("firebrick", alpha.f = 0.8)

# Plot the original and fitted values using base R
transparent_firebrick <- adjustcolor("firebrick", alpha.f = 0.6)
plot_fitted_base <- function(df, title) {
  plot(df$Time, df$Original, type = "l", col = "black", lwd = 2, ylab = "Value", xlab = "Time", main = title)
  lines(df$Time, df$Fitted, col = transparent_firebrick, lwd = 2)
  legend("topright", legend = c("Original", "Fitted"), col = c("black", transparent_firebrick), lwd = 2)
}

# Plot each dataset
par(mfrow = c(1,1))  # Set up a 2x2 plotting space
plot_fitted_base(df_fitted_DA_CH, "Day Ahead Price CH")
plot_fitted_base(df_fitted_DA_DE, "Day Ahead Price DE")
plot_fitted_base(df_fitted_AUCT_CH_DE, "Auction Price CH-DE")
plot_fitted_base(df_fitted_AUCT_DE_CH, "Auction Price DE-CH")
```


 
#Swiss DA price
```{r}
df_scores= data.frame(date= df_CH_DA_test$date[1:743], .pred= predictions_DA_CH, actual=df_CH_DA_test$day_ahead_price_ch[1:743] )

# Assuming 'df' is your dataframe
write.csv(df_scores, "/Users/olivier/Documents/ETH 2. Semester¨/Statslab/Scores Std/1 - GARCH - Prediction - CH - Data Predictive/holdout_predictions.csv", row.names = FALSE)


print("saved")
```

#German DA price
```{r}
df_scores= data.frame(date= df_DE_DA_test$date[1:743], .pred= predictions_DA_DE, actual=df_DE_DA_test$day_ahead_price_de[1:743] )

# Assuming 'df' is your dataframe
write.csv(df_scores, "/Users/olivier/Documents/ETH 2. Semester¨/Statslab/Scores Std/1 - GARCH - Prediction - DE - Data Predictive/holdout_predictions.csv", row.names = FALSE)


print("saved")
```

#Auciton CH_DE price
```{r}
df_scores= data.frame(date= df_CH_DE_AUCT_test$date[1:743], .pred= predictions_AUCT_CH_DE, actual=df_CH_DE_AUCT_test$auction_price_ch_de[1:743] )

# Assuming 'df' is your dataframe
write.csv(df_scores, "/Users/olivier/Documents/ETH 2. Semester¨/Statslab/Scores Std/1 - GARCH - Prediction - CHDE - Data Predictive/holdout_predictions.csv", row.names = FALSE)


print("saved")
```

#Auction DE_CH
```{r}
df_scores= data.frame(date= df_DE_CH_AUCT_test$date[1:743], .pred= predictions_AUCT_DE_CH, actual=df_CH_DE_AUCT_test$auction_price_de_ch[1:743] )

# Assuming 'df' is your dataframe
write.csv(df_scores, "/Users/olivier/Documents/ETH 2. Semester¨/Statslab/Scores Std/1 - GARCH - Prediction - DECH - Data Predictive/holdout_predictions.csv", row.names = FALSE)


print("saved")
```

```{r}
#save(predictions_DA_CH,predictions_DA_DE, predictions_AUCT_CH_DE,predictions_AUCT_DE_CH,  fit_DA_CH_ARMA_GARCH, fit_DA_DE_ARMA_GARCH, fit_AUCT_CH_DE_ARMA_GARCH, fit_AUCT_DE_CH_ARMA_GARCH, file = "ALL CORRECT w. retrining (ugarch) correct predictors, (predictive).RData") 
#print("saved")

load("ALL CORRECT w. retrining (ugarch) correct predictors, (predictive).RData")
```



```{r}
save(predictions_DA_CH, fit_DA_CH_ARMA_GARCH, file = "Correct model + predictions Swiss Day ahead price, correct predictors, (predictive).RData") 

load("Correct model + predictions Swiss Day ahead price, correct predictors, (predictive).RData")
```

```{r}
save(predictions_DA_DE, fit_DA_DE_ARMA_GARCH, file = "Correct model + predictions German Day ahead price, correct predictors, (predictive).RData") 

load("Correct model + predictions German Day ahead price, correct predictors, (predictive).RData") 
```

```{r}
save(predictions_AUCT_DE_CH, fit_AUCT_CH_DE_ARMA_GARCH, file = "Correct model + predictions Auction price CH_DE, correct predictors, (predictive).RData") 

load("Correct model + predictions Auction price CH_DE, correct predictors, (predictive).RData") 
```

```{r}
save(predictions_AUCT_DE_CH, fit_AUCT_DE_CH_ARMA_GARCH, file = "Correct model + predictions Auction price DE_CH, correct predictors, (predictive).RData") 

load("Correct model + predictions Auction price DE_CH, correct predictors, (predictive).RData")
```

#Save Swiss DA predictions
```{r}
df_test= df_CH_DA_test
forecast_values =predictions_DA_CH
response_test= df_CH_DA_test$day_ahead_price_ch

df_scores= data.frame(date= df_test$date[1:743], .pred= forecast_values[1:743], actual=response_test[1:743] )

write.csv(df_scores, "/Users/olivier/Documents/ETH 2. Semester¨/Statslab/Scores DA prices/1 - GARCH - Prediction - CH - Data Predictive/holdout_predictions.csv", row.names = FALSE)

print("saved")
```

German DA predictions
```{r}
df_test= df_DE_DA_test
forecast_values =predictions_DA_DE
response_test= df_CH_DA_test$day_ahead_price_de

df_scores= data.frame(date= df_test$date[1:743], .pred= forecast_values[1:743], actual=response_test[1:743] )

write.csv(df_scores, "/Users/olivier/Documents/ETH 2. Semester¨/Statslab/Scores DA prices/1 - GARCH - Prediction - DE - Data Predictive/holdout_predictions.csv", row.names = FALSE)

print("saved")
```
Auction DE-CH predictions
```{r}
df_test= df_DE_CH_AUCT_test
forecast_values =predictions_AUCT_DE_CH
response_test= df_CH_DA_test$auction_price_de_ch

df_scores= data.frame(date= df_test$date[1:743], .pred= forecast_values[1:743], actual=response_test[1:743] )

write.csv(df_scores, "/Users/olivier/Documents/ETH 2. Semester¨/Statslab/Scores Auction prices/1 - GARCH - Prediction - DECH - Data Predictive/holdout_predictions.csv", row.names = FALSE)

print("saved")
```

Auction CH_DE predictions
```{r}
df_test= df_CH_DE_AUCT_test
forecast_values =predictions_AUCT_CH_DE
response_test= df_CH_DA_test$auction_price_ch_de

df_scores= data.frame(date= df_test$date[1:743], .pred= forecast_values[1:743], actual=response_test[1:743] )

write.csv(df_scores, "/Users/olivier/Documents/ETH 2. Semester¨/Statslab/Scores Auction prices/1 - GARCH - Prediction - CHDE - Data Predictive/holdout_predictions.csv", row.names = FALSE)

print("saved")
```

Visualizion OOS fit
```{r}
#Visualization of out of sample fit:
transparent_firebrick <- adjustcolor("firebrick", alpha.f = 0.8)
plot(df_CH_DA_test$day_ahead_price_ch[0:743], type="l", main="Swiss Day Ahead Prices - Predictions",  xlab = " ", ylab = "Spot Price", xaxt = "n", ylim=c(0,180), lwd=1.5)
lines(predictions_DA_CH, col=transparent_firebrick, lwd=1.5)
axis(1, at = seq(1, 743, by = 100), labels = df_DE_DA_test$date[seq(1, 743, by = 100)], las = 2, cex.axis = 0.7)
legend("topright", legend = c("True Prices", "Predictions: ARMA(2,2)-GARCH(1,1)"), col = c("black", transparent_firebrick), lwd = 2, cex=0.7)

plot(df_DE_DA_test$day_ahead_price_de[0:743], type="l", main="German Day Ahead Prices - Predictions", ylim=c(-50,220), xlab = " ", ylab = "Price", xaxt = "n", , lwd=1.5,  xaxt = "n")
lines(predictions_DA_DE, col=transparent_firebrick, lwd=1.5)
axis(1, at = seq(1, 743, by = 100), labels = df_DE_DA_test$date[seq(1, 743, by = 100)], las = 2, cex.axis = 0.7)
legend("topright", legend = c("True Prices", "Predictions: ARMA(2,2)-GARCH(1,1)"), col = c("black", transparent_firebrick), lwd = 2, cex=0.7)

plot(df_DE_CH_AUCT_test$auction_price_de_ch[0:743], type="l", main="Auction Prices DE_CH - Predictions", xlab = " ", ylab = "Price", xaxt = "n", ylim= c(0,80))
lines(pmax(0,predictions_AUCT_DE_CH), col=transparent_firebrick)
axis(1, at = seq(1, 743, by = 360), labels = df_DE_DA_test$date[seq(1, 743, by = 360)], las = 2, cex.axis = 0.7)
legend("topright", legend = c("True Prices", "Predictions: ARMA(2,2)-GARCH(1,1)"), col = c("black",transparent_firebrick), lwd = 2, cex=0.7)

plot(df_CH_DE_AUCT_test$auction_price_ch_de[0:743], type="l", main="Auction Prices CH_DE - Predictions", xlab = " ", ylab = "Price", xaxt = "n", ylim= c(0,100))
lines(pmax(0,predictions_AUCT_CH_DE), col=transparent_firebrick)
axis(1, at = seq(1, 8734, by = 360), labels = df_DE_DA_train$date[seq(1, 8734, by = 360)], las = 2, cex.axis = 0.7)

legend("topright", legend = c("True Prices", "Predictions: ARMA(2,2)-GARCH(1,1)"), col = c("black",transparent_firebrick), lwd = 2, cex=0.7)
```


```{r}
library(yardstick)

response_test = df_CH_DA_test$day_ahead_price_ch[0:743]
forecast_values= predictions_DA_CH
rsquared= rsq_vec(as.numeric(response_test),as.numeric(forecast_values))
rmse=  sqrt(mean((response_test -forecast_values)^2))
MAE= mean(response_test -forecast_values)
```



Visualization of fit
```{r}

#Visualization of fit:
# Extract the fitted values
fitted_values_DA_CH <- sinh(fitted(fit_DA_CH_ARMA_GARCH))

# Combine the original and fitted values into data frames
df_fitted_DA_CH <- data.frame( Original = sinh(train_data_CH_DA$asinh_price_ch[1:720]), Fitted = fitted_values_DA_CH[1:720])

transparent_firebrick <- adjustcolor("firebrick", alpha.f = 0.8)
# Plot the original and fitted values using base R
plot_fitted_base <- function(df, title) {
  plot( df$Original, type = "l", col = "black", lwd = 1, ylab = "Value", main = title, xlab="", xaxt = "n")
  lines(df$Fitted, col = transparent_firebrick, lwd = 1)
  legend("bottomright", legend = c("True Prices", "Fitted: ARMA(3,3)-GARCH(3,3)"), col = c("black", "firebrick"), lwd = 2, cex=0.8)
}

# Plot each dataset
par(mfrow = c(1,1))  # Set up a 2x2 plotting space
plot_fitted_base(df_fitted_DA_CH, "Day Ahead Price CH")
axis(1, at = seq(1, 743, by = 100), labels = df_DE_DA_train$date[seq(1, 743, by = 100)], las = 2, cex.axis = 0.7)

```



###Resiudal Analysis:###
```{r}
library(Dowd)
# Plotting the residuals
residuals_DA_price_ch  <- residuals(fit_DA_CH_ARMA_GARCH)



# Plot histogram of residuals
hist(residuals_DA_price_ch, breaks = 20, main = "Histogram of Residuals")


# Plot QQ plot of residuals
residuals_DA_CH_asinh= fitted(fit_DA_CH_ARMA_GARCH) - c(asinh(df_CH_DA_train$day_ahead_price_ch), asinh(df_CH_DA_test$day_ahead_price_ch))[1:length(fitted(fit_DA_CH_ARMA_GARCH))]

sigmas_DA_CH= sigma(fit_DA_CH_ARMA_GARCH)

epsilons= residuals_DA_CH_asinh/sigmas_DA_CH
d_o_F_of_T_dist= 3.465331
TQQPlot(as.vector(residuals_DA_price_ch),d_o_F_of_T_dist)



```

Tukey Anscombe plot
```{r}
# Required Libraries
library(rugarch)
library(ggplot2)



# Function to create Tukey-Anscombe plot
create_tukey_anscombe_plot <- function(fit_model, model_name) {
  fitted_values <- sinh(fitted(fit_model))
  residuals <- residuals(fit_model, standardize = TRUE)  # standardized residuals

  # Create a data frame for plotting
  plot_data <- data.frame(Fitted = fitted_values, Residuals = residuals)
  
  # Create the plot using ggplot2
  plot <- ggplot(plot_data, aes(x = Fitted, y = Residuals)) +
    geom_point() +
    geom_hline(yintercept = 0, linetype = "dashed", color = "red") +
    geom_smooth(method = "loess", color = "red", se = FALSE) +
    theme_minimal() +
    labs(title = paste("Tukey-Anscombe Plot for", model_name),
         x = "Fitted Values",
         y = "Standardized Residuals")
  
  print(plot)
}

# Create Tukey-Anscombe plots for each model
create_tukey_anscombe_plot(fit_DA_CH_ARMA_GARCH, "Day Ahead Price CH")
create_tukey_anscombe_plot(fit_DA_DE_ARMA_GARCH, "Day Ahead Price DE")
create_tukey_anscombe_plot(fit_AUCT_DE_CH_ARMA_GARCH, "Auction Price DE_CH")
create_tukey_anscombe_plot(fit_AUCT_CH_DE_ARMA_GARCH, "Auction price CH_DE")

```

Deleting outliers
```{r}
# Required Libraries
library(rugarch)
library(ggplot2)

# Function to create Tukey-Anscombe plot
create_tukey_anscombe_plot <- function(fit_model, model_name, lower_threshold = -500, upper_threshold = 500) {
  fitted_values <- sinh(fitted(fit_model))
  residuals <- residuals(fit_model, standardize = TRUE)  # standardized residuals

  # Create a data frame for plotting
  plot_data <- data.frame(Fitted = fitted_values, Residuals = residuals)
  
  # Filter out the outliers based on the thresholds
  plot_data_filtered <- plot_data[plot_data$Fitted > lower_threshold & plot_data$Fitted < upper_threshold, ]
  
  # Create the plot using ggplot2
  plot <- ggplot(plot_data_filtered, aes(x = Fitted, y = Residuals)) +
    geom_point() +
    geom_hline(yintercept = 0, linetype = "dashed", color = "red") +
    geom_smooth(method = "loess", color = "red", se = FALSE) +
    theme_minimal() +
    labs(title = paste("Tukey-Anscombe Plot for", model_name),
         x = "Fitted Values",
         y = "Standardized Residuals")
  
  print(plot)
}

# Create Tukey-Anscombe plots for each model with thresholds
create_tukey_anscombe_plot(fit_DA_CH_ARMA_GARCH, "Day Ahead Price CH", lower_threshold = -500, upper_threshold = 500)
create_tukey_anscombe_plot(fit_DA_DE_ARMA_GARCH, "Day Ahead Price DE", lower_threshold = -500, upper_threshold = 500)
```

```{r}
# Required Libraries
library(rugarch)

# Apply the sinh transformation to the fitted values
fitted_values_CH_DA <- sinh(fitted(fit_DA_CH_ARMA_GARCH))
fitted_values_DE_DA <- sinh(fitted(fit_DA_DE_ARMA_GARCH))

# Filter the fitted values to only include those below 500
filtered_fitted_values_CH_DA <- fitted_values_CH_DA[fitted_values_CH_DA < 500 & fitted_values_CH_DA >-500]
filtered_fitted_values_DE_DA <- fitted_values_DE_DA[fitted_values_DE_DA < 500 & fitted_values_DE_DA >-500]

# Display the filtered fitted values
hist(filtered_fitted_values_CH_DA, breaks=20, main= "Fitted values Swiss DA prices")
hist(filtered_fitted_values_DE_DA, breaks=20,  main= "Fitted values German DA prices")

#No need to filter values of. Auction price
hist(sinh(fitted(fit_AUCT_DE_CH_ARMA_GARCH)), breaks=40, main= "Fitted values Auction price DE_CH")
hist(sinh(fitted(fit_AUCT_CH_DE_ARMA_GARCH)), breaks=40,  main= "Fitted values Auction price CH_DE")

```
plotting residuals vs time whole 2023
```{r}
library(lubridate)

# Generate a sequence of daily dates for the year 2023
daily_dates_2023 <- seq(ymd("2023-01-01"), ymd("2023-12-31"), by = "day")

# Repeat each date 24 times to represent each hour in a day
repeated_daily_dates_2023 <- rep(daily_dates_2023, each = 24)

# Function to create a residuals dataframe and plot
plot_residuals <- function(model, main_title) {
  # Assuming each residuals vector has 8734 entries (24 hours * 365 days)
  residuals_df <- data.frame(Date = repeated_daily_dates_2023[1:8734], Residuals = residuals(model)[1:8734])

  transparent_firebrick <- adjustcolor("firebrick", alpha.f = 0.8)

  # Plot with adjusted x-axis
  plot(sinh(residuals_df$Residuals), type="l", main=main_title, 
       xlab = "2023", ylab = "Spot Price", xaxt = "n", lwd=0.5)

  # Add axis with monthly labels
  axis(1, at = seq(1, 8734, length.out = 12), 
       labels = format(seq(ymd("2023-01-01"), ymd("2023-12-01"), by = "month"), "%b"), 
       las = 2, cex.axis = 0.7)
}

# Plot for fit_AUCT_CH_DE_ARMA_GARCH
plot_residuals(fit_AUCT_CH_DE_ARMA_GARCH, "Auction price CH_DE")

# Plot for fit_DA_CH_ARMA_GARCH
plot_residuals(fit_DA_CH_ARMA_GARCH, "Swiss Day Ahead Prices ")

# Plot for fit_DA_DE_ARMA_GARCH
plot_residuals(fit_DA_DE_ARMA_GARCH, "German Day Ahead Prices")

# Plot for fit_AUCT_DE_CH_ARMA_GARCH
plot_residuals(fit_AUCT_DE_CH_ARMA_GARCH, "Auction price DE_CH")

```


plotting residuals vs time February 2023
```{r}
library(lubridate)

# Generate a sequence of daily dates for the year 2023
daily_dates_2023 <- seq(ymd("2023-01-01"), ymd("2023-12-31"), by = "day")

# Repeat each date 24 times to represent each hour in a day
repeated_daily_dates_2023 <- rep(daily_dates_2023, each = 24)

# Generate a sequence of dates for January 2023
january_dates_2023 <- seq(ymd("2023-01-01"), ymd("2023-01-31"), by = "days")

# Function to create a residuals dataframe and plot
plot_residuals <- function(model, main_title) {
  # Assuming each residuals vector has 744 entries (24 hours * 31 days)
  residuals_df <- data.frame(Date = repeated_daily_dates_2023[744:1416], Residuals = residuals(model)[744:1416])

  transparent_firebrick <- adjustcolor("firebrick", alpha.f = 0.8)

  # Plot with adjusted x-axis
  plot(sinh(residuals_df$Residuals), type="l", main=main_title, 
       xlab = "February 2023", ylab = "Spot Price", xaxt = "n", lwd=0.5)

  # Add axis with daily labels for January, every 3rd day
  axis(1, at = seq(1, 672, by = 72), 
       labels = format(seq(ymd("2023-02-01"), ymd("2023-02-28"), by = "3 days"), "%d-%b"), 
       las = 2, cex.axis = 0.7)
}

# Plot for fit_AUCT_CH_DE_ARMA_GARCH
plot_residuals(fit_AUCT_CH_DE_ARMA_GARCH, "Auction price CH_DE")

# Plot for fit_DA_CH_ARMA_GARCH
plot_residuals(fit_DA_CH_ARMA_GARCH, "Swiss Day Ahead Prices")

# Plot for fit_DA_DE_ARMA_GARCH
plot_residuals(fit_DA_DE_ARMA_GARCH, "German Day Ahead Prices")

# Plot for fit_AUCT_DE_CH_ARMA_GARCH
plot_residuals(fit_AUCT_DE_CH_ARMA_GARCH, "Auction price DE_CH")

```

ACF plot
```{r}
library(forecast)
library(ggplot2)

# Function to plot ACF of residuals with a maximum of 48 lags using ggplot2
plot_acf_residuals_ggplot <- function(model, title, max_lag = 60) {
  # Extract residuals from the model and convert to a numeric vector
  residuals_xts <- residuals(model)
  residuals <- as.numeric(residuals_xts)
  
  # Compute the ACF values
  acf_res <- acf(residuals, lag.max = max_lag, plot = FALSE)
  
  # Prepare data for ggplot
  acf_data <- data.frame(Lag = acf_res$lag, ACF = acf_res$acf)
  
  # Create the ACF plot using ggplot2
  ggplot(acf_data, aes(x = Lag, y = ACF)) +
    geom_hline(yintercept = 0, color = "black") +
    geom_segment(aes(xend = Lag, yend = 0)) +
    geom_hline(yintercept = c(-1.96/sqrt(length(residuals)), 1.96/sqrt(length(residuals))), linetype = "dashed", color = "blue") +
    labs(title = paste("ACF of Residuals:", title), x = "Lag", y = "ACF") +
    theme_minimal()
}

# Assuming fit_AUCT_CH_DE_ARMA_GARCH, fit_DA_CH_ARMA_GARCH,
# fit_DA_DE_ARMA_GARCH, and fit_AUCT_DE_CH_ARMA_GARCH are your models

# Plot for fit_AUCT_CH_DE_ARMA_GARCH
plot_acf_residuals_ggplot(fit_AUCT_CH_DE_ARMA_GARCH, "Auction price CH_DE")

# Plot for fit_DA_CH_ARMA_GARCH
plot_acf_residuals_ggplot(fit_DA_CH_ARMA_GARCH, "Swiss Day Ahead Prices")

# Plot for fit_DA_DE_ARMA_GARCH
plot_acf_residuals_ggplot(fit_DA_DE_ARMA_GARCH, "German Day Ahead Prices")

# Plot for fit_AUCT_DE_CH_ARMA_GARCH
plot_acf_residuals_ggplot(fit_AUCT_DE_CH_ARMA_GARCH, "Auction price DE_CH")


```

