---
title: "fastTS_precition _de_de_ch"
author: "Federico Deotto"
date: "2024-05-10"
output: html_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(lubridate)
library(ggsci)
library(scales)
library(shiny)
library(ncvreg)

library(fastTS)
library(xts)




setwd(dirname(rstudioapi::getActiveDocumentContext()$path))
source('builtin_functions.R')

# Default theme for charts
theme_set(
  theme_bw() +
    theme(  
      plot.title = element_text(face = "bold", size = 14),
      plot.subtitle = element_text(
        face = "italic", size = 10, colour = "green50"
      )
    )
)
```

## Goal

In this Rmarkdown, we analyze the prices using a prediction perspective, i.e., we try to use only the information that is available before the JAO auctions in the morning. Since the time series over the entire time window exhibits clear non-stationarity, we decided to focus on the last period that seems to present 'constant' volatility and is more relevant to our client. We will explore different transformations of the responses and also alter the feature space.


# JAO Price de_ch
```{r}
time_cut <- as.POSIXct("2023-03-01 00:00:00", tz = "UTC")

df <- read_csv("../00 Data Retrieval and Cleaning/0_df_final_imputed_shifting_a_de_ch.csv") %>% 
  filter(date >= time_cut)

#Remove variables with constant variance; they are redundant since we have cut the time window (consider for example year dummies)

variances = df %>%  
  select(-c(date)) %>%  
  mutate_all(as.numeric) %>%
  apply(2, var)

null_variance_cols <- names(variances[variances == 0])

df  %>% 
  nrow() * 0.7

max_lag =24*30 # 1 month - this is the amount of data that we will lose by shifting
```

We can observe that the log transformation is more suitable for the modeling phase. Additionally, we need to shift the data since we have null prices.
```{r}
df %>% 
  ggplot(aes(date))+
  geom_line(aes(y = auction_price_de_ch))+
  labs(title = 'Price original', x = 'Date', y = 'Price')

df %>% 
  mutate(auction_price_de_ch = log(auction_price_de_ch - min(auction_price_de_ch) +1)) %>% 
  ggplot(aes(date))+
  geom_line(aes(y = auction_price_de_ch))+
  labs(title = 'Price log', x = 'Date', y = 'Price')

df %>% 
  mutate(auction_price_de_ch = tanh(auction_price_de_ch)) %>% 
  ggplot(aes(date))+
  geom_line(aes(y = auction_price_de_ch))+
  labs(title = 'Price tanh', x = 'Date', y = 'Price')
```

# Modelling swiss price:

We will adapt two kinds of models:

1)Utilize lags ranging from 1 to 24. This information is not available before the auction, but once it's time for prediction, we can use the prediction of the previous lag, accumulating uncertainty.
2)Adapt a model without using these lags of the response.

For both model we will also consider the squared version  of those coefficient that where significant in the best model based on the AICc statistic.


```{r}
models1 <- df %>% 
  mutate(auction_price_de_ch = log(auction_price_de_ch  + 1)) %>% 
  model_list(p.train = 0.7, response = 'auction_price_de_ch', nlag = max_lag)

summary(models1)

temp <- summary(models1)$table
variables_to_square <- temp %>%
  rownames() %>%
  enframe(name = NULL) %>%
  filter(!str_detect(value, 'lag') & !str_detect(value, 'cal')) %>% 
  pull() %>% 
  sort()

rm(temp)

models2 <- df %>% 
  mutate(auction_price_de_ch = log(auction_price_de_ch  + 1)) %>%
  mutate(across(variables_to_square, ~ .^2, .names = "{.col}_squared")) %>% 
  model_list(p.train = 0.7, response = 'auction_price_de_ch', nlag = max_lag)

summary(models2)

temp <- summary(models2)$table
variables_after_square <- temp %>%
  rownames() %>%
  enframe(name = NULL) %>%
  filter(!str_detect(value, 'lag') & !str_detect(value, 'cal')) %>% 
  pull() %>% 
  sort()

rm(temp)

print('Element present in both models')
intersect(variables_to_square, variables_after_square)

print('Element present only in first models, and not relevant anymore')
setdiff(variables_to_square, variables_after_square)

print('Element present only in second models')
setdiff(variables_after_square, variables_to_square)
```


Second type of models

```{r}
models3 <- df %>% 
  mutate(auction_price_de_ch = log(auction_price_de_ch  + 1)) %>% 
  model_list_prediction(ptrain = 0.7, response = 'auction_price_de_ch', n_lags_max = max_lag)

summary(models3)

temp <- summary(models3)$table
variables_to_square <- temp %>%
  rownames() %>%
  enframe(name = NULL) %>%
  filter(!str_detect(value, 'lag') & !str_detect(value, 'cal')) %>% 
  pull() %>% 
  sort()

rm(temp)

models4 <- df %>% 
  mutate(auction_price_de_ch = log(auction_price_de_ch  + 1)) %>%
  mutate(across(variables_to_square, ~ .^2, .names = "{.col}_squared")) %>% 
  model_list_prediction(ptrain = 0.7, response = 'auction_price_de_ch', n_lags_max = max_lag)

summary(models4)

temp <- summary(models4)$table
variables_after_square <- temp %>%
  rownames() %>%
  enframe(name = NULL) %>%
  filter(!str_detect(value, 'lag') & !str_detect(value, 'cal')) %>% 
  pull() %>% 
  sort()

rm(temp)
print('Element present in both models')
intersect(variables_to_square, variables_after_square)

print('Element present only in first models, and not relevant anymore')
setdiff(variables_to_square, variables_after_square)

print('Element present only in second models')
setdiff(variables_after_square, variables_to_square)
```

#Analisys of results.

```{r}
diagnostic1 <- model_selected(models1, lag.max = max_lag, p.train = 0.7)
prediction1 <- prediction_step(models1)
jao_predictions(prediction1)
```

```{r}
diagnostic2 <- model_selected(models2, lag.max = max_lag, p.train = 0.7)
prediction2 <- prediction_step(models2)
jao_predictions(prediction2)
```

```{r}
diagnostic3 <- model_selected(models3, lag.max = max_lag, p.train = 0.7)
prediction3 <- prediction_model(models3)
jao_predictions(prediction3)
```

```{r}
diagnostic4 <- model_selected(models4, lag.max = max_lag, p.train = 0.7)
prediction4 <- prediction_model(models4)
jao_predictions(prediction4)
```


The best model is the second type (no lags) with the squared version of the variables.

#  Seond type of dataframe - future observations. NO MORE PREDICTIVE

# Swiss Price
```{r}
df <- read_csv("../00 Data Retrieval and Cleaning/0_df_final_imputed_shifting_c_de_ch.csv") %>% 
  filter(date >= time_cut)

#Remove variables with constant variance; they are redundant since we have cut the time window (consider for example year dummies)

variances = df %>%  
  select(-c(date)) %>%  
  mutate_all(as.numeric) %>%
  apply(2, var)

null_variance_cols <- names(variances[variances == 0])

df  %>% 
  nrow() * 0.7

max_lag =24*30 # 1 month - this is the amount of data that we will lose by shifting
```

# Modelling german price:

As we did previously we will adapt two kinds of models.


```{r}
models1 <- df %>% 
  mutate(auction_price_de_ch = log(auction_price_de_ch  + 1)) %>% 
  model_list(p.train = 0.7, response = 'auction_price_de_ch', nlag = max_lag)

summary(models1)

temp <- summary(models1)$table
variables_to_square <- temp %>%
  rownames() %>%
  enframe(name = NULL) %>%
  filter(!str_detect(value, 'lag') & !str_detect(value, 'cal')) %>% 
  pull() %>% 
  sort()

rm(temp)

models2 <- df %>% 
  mutate(auction_price_de_ch = log(auction_price_de_ch  + 1)) %>%
  mutate(across(variables_to_square, ~ .^2, .names = "{.col}_squared")) %>% 
  model_list(p.train = 0.7, response = 'auction_price_de_ch', nlag = max_lag)

summary(models2)

temp <- summary(models2)$table
variables_after_square <- temp %>%
  rownames() %>%
  enframe(name = NULL) %>%
  filter(!str_detect(value, 'lag') & !str_detect(value, 'cal')) %>% 
  pull() %>% 
  sort()

rm(temp)

print('Element present in both models')
intersect(variables_to_square, variables_after_square)

print('Element present only in first models, and not relevant anymore')
setdiff(variables_to_square, variables_after_square)

print('Element present only in second models')
setdiff(variables_after_square, variables_to_square)
```


Second type of models

```{r}
models3 <- df %>% 
  mutate(auction_price_de_ch = log(auction_price_de_ch  + 1)) %>% 
  model_list_prediction(ptrain = 0.7, response = 'auction_price_de_ch', n_lags_max = max_lag)

summary(models3)

temp <- summary(models3)$table
variables_to_square <- temp %>%
  rownames() %>%
  enframe(name = NULL) %>%
  filter(!str_detect(value, 'lag') & !str_detect(value, 'cal')) %>% 
  pull() %>% 
  sort()

rm(temp)

models4 <- df %>% 
  mutate(auction_price_de_ch = log(auction_price_de_ch  + 1)) %>%
  mutate(across(variables_to_square, ~ .^2, .names = "{.col}_squared")) %>% 
  model_list_prediction(ptrain = 0.7, response = 'auction_price_de_ch', n_lags_max = max_lag)

summary(models4)

temp <- summary(models4)$table
variables_after_square <- temp %>%
  rownames() %>%
  enframe(name = NULL) %>%
  filter(!str_detect(value, 'lag') & !str_detect(value, 'cal')) %>% 
  pull() %>% 
  sort()

rm(temp)
print('Element present in both models')
intersect(variables_to_square, variables_after_square)

print('Element present only in first models, and not relevant anymore')
setdiff(variables_to_square, variables_after_square)

print('Element present only in second models')
setdiff(variables_after_square, variables_to_square)
```

#Analisys of results.

```{r}
diagnostic1 <- model_selected(models1, lag.max = max_lag, p.train = 0.7)
prediction1 <- prediction_step(models1)
jao_predictions(prediction1)
```

```{r}
diagnostic2 <- model_selected(models2, lag.max = max_lag, p.train = 0.7)
prediction2 <- prediction_step(models2)
jao_predictions(prediction2)
```

```{r}
diagnostic3 <- model_selected(models3, lag.max = max_lag, p.train = 0.7)
prediction3 <- prediction_model(models3)
jao_predictions(prediction3)
```

```{r}
diagnostic4 <- model_selected(models4, lag.max = max_lag, p.train = 0.7)
prediction4 <- prediction_model(models4)
jao_predictions(prediction4)
```


